---
title: "Group Project 2"
author: "Ian"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readr)
#read in data
confirmed_global <- read.csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv")

confirmed_us <- read.csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_US.csv")

deaths_global <- read.csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv")

deaths_us <- read.csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_US.csv")
```
Objective 1 - Global Map
Data Wrangling
```{r}
library(dplyr)

#change column name of last cumulative column to "Confirmed_Cases"
colnames(confirmed_global)[ncol(confirmed_global)] = "Confirmed_Cases"
#group table by country and calculate average lat and long for countries and sum of confirmed cases
grouped_confirmed <- confirmed_global %>% 
  select(1:4, ncol(confirmed_global)) %>% 
      group_by(Country.Region) %>%
          summarize(mean_lat=mean(Lat),
                    mean_long=mean(Long),
                    sum(Confirmed_Cases))
#change column name of last cumulative column to "Deaths"
colnames(deaths_global)[ncol(deaths_global)] = "Deaths"
#group table by country and calculate average lat and long for countries and sum of deaths
grouped_deaths <- deaths_global %>%
  select(1:4, ncol(deaths_global)) %>%
    group_by(Country.Region) %>%
      summarize(mean_lat=mean(Lat),
                mean_long=mean(Long),
                sum(Deaths))
#join the 2 data frames and remove the duplicate lat and long columns
global_stats <- merge(x=grouped_confirmed, y=grouped_deaths, by="Country.Region") %>% select(-5, -6)

```
Map Making
```{r}
library("leaflet")
global_stats$`sum(Confirmed_Cases)` = as.factor(global_stats$`sum(Confirmed_Cases)`)
pal <- colorFactor("RdYlGn", global_stats$`sum(Confirmed_Cases)`, reverse = TRUE)


leaflet(data = global_stats) %>%
  addProviderTiles("CartoDB.Positron") %>%
  setView(lng = -100, lat = 40, zoom = 5) %>%
  addCircles(
    lat = ~mean_lat.x,
    lng = ~mean_long.x,
    popup = paste("Country:", global_stats$Country.Region, "<br>",
                  "Confirmed Cases: ", global_stats$`sum(Confirmed_Cases)`, "<br>",
                  "Deaths:" , global_stats$`sum(Deaths)`),
    radius = 500,
    color = ~pal(global_stats$`sum(Confirmed_Cases)`),
    fillOpacity = .5)
  
  
  
